Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-11 method.tex
Minor tweaks in section 2.3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1026 ccdd05d5-56c1-461a-bf98-85f6876523b4

The first step of our proposed method is to divide the templates into time
slices in a \TD{} analogue to the \FD{} decomposition described
in \citet{Marion2004, Buskulic2010}.  To wit, we decompose each template
$h_{i}[k]$ into a sum of $S$ non-overlapping filters
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\begin{cases}
		h_i^s[k] & \textrm{if } t^s \leqslant k / f^0 < t^{s+1} \\
		0 & \textrm{otherwise}
	\end{cases}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
t^S = N$.  A matched filter is constructed for each time slice.  The outputs
form an ensemble of partial \SNR{} streams.  By linearity of the filtering
process, these partial \SNR\ streams can be suitably time-delayed and summed to
 reproduce the \SNR\ of the full template.

Since waveforms with neighboring intrinsic source parameters $\mathbf\Theta$
 have similar time-frequency evolution, it is possible to design computationally
efficient time slices for an extended region of parameter space rather than to
design different time slices for each template.
For concreteness and simplicity, consider an inspiral waveform in the
quadrapole approximation, for which the time-frequency relation is
%
\begin{equation} \label{eq:fgw}
%
f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
\right]^{3/8}.
%
\end{equation}
%
Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
M_\odot / c^3 \approx 5 \upmu\mathrm{s}$) and $t$ is the time relative to the
coalescence of the binary~\citep{findchirppaper, kidder1992}.
The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us
to choose time-slice boundaries that require substantially less bandwidth at
early times in the inspiral.

In any given interval of time, the filters are band-limited across the enitre
template bank. Usually the template is truncated at some prescribed time $t^0$,
or equivalently frequency $f_\mathrm{hi}$, often chosen to correspond to the
\ISCO. An inspiral signal will enter the detection band at some low frequency,
$f = f_\mathrm{low}$, corresponding to a time $t_\mathrm{low}$.  The end of the
template is then critically sampled at a rate of $2 f_\mathrm{hi}$ but the 
beginning of the template is critically sampled at $2 f_\mathrm{lo}$. In any
time interval smaller than the duration of the template, the bandwidth of the
filters across the entire template bank may be significantly less and
a lower sampling rate may be used for all filters without aliasing.

>>>>

The first step of our proposed method is to divide the templates into time
slices in a \TD{} analogue to the \FD{} decomposition described
in \citet{Marion2004, Buskulic2010}.  We decompose each template
$h_{i}[k]$ into a sum of $S$ non-overlapping templates
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\begin{cases}
		h_i^s[k] & \textrm{if } t^s \leqslant k / f^0 < t^{s+1} \\
		0 & \textrm{otherwise}
	\end{cases}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
t^S = N$.  The outputs of these new time-slice filters
form an ensemble of partial \SNR{} streams.  By linearity of the filtering
process, these partial \SNR\ streams can be suitably summed to reproduce the
\SNR\ of the full template.

Since waveforms with neighboring intrinsic source parameters $\mathbf\Theta$
 have similar time-frequency evolution, it is possible to design computationally
efficient time slices for an extended region of parameter space rather than to
design different time slices for each template.

For concreteness and simplicity, consider an inspiral waveform in the
quadrapole approximation, for which the time-frequency relation is
%
\begin{equation} \label{eq:fgw}
%
f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
\right]^{3/8}.
%
\end{equation}
%
Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
M_\odot / c^3 \approx 5 \upmu\mathrm{s}$) and $t$ is the time relative to the
coalescence of the binary~\citep{findchirppaper, kidder1992}.
This monotonic time-frequency relationship allows us
to choose time-slice boundaries that require substantially less bandwidth at
early times in the inspiral.

In any given interval of time, the filters are band-limited across the enitre
template bank. Usually the template is truncated at some prescribed time $t^0$,
or equivalently frequency $f_\mathrm{hi}$, often chosen to correspond to the
\ISCO. An inspiral signal will enter the detection band at some low frequency,
$f = f_\mathrm{low}$, corresponding to a time $t_\mathrm{low}$.  The end of the
template is then critically sampled at a rate of $2 f_\mathrm{hi}$ but the 
beginning of the template is critically sampled at $2 f_\mathrm{low}$. In any
time interval smaller than the duration of the template, the bandwidth of the
filters across the entire template bank may be significantly less and
a lower sampling rate may be used for all filters without aliasing.
